### 2017-02-15 Lab_1/acro.tex
Filled in first three questions of lab 1.

%!TEX root =labo.tex

\begin{acronym}
\acro{aes}[AES]{Advanced Encryption Standard}
\acro{ap}[AP]{access point}
\acro{arp}[ARP]{Address Resolution Protocol}
\acro{bipt}[BIPT]{Belgisch Instituut voor Postdiensten en Telecommunicatie - Belgian Institute for Postal services and Telecommunications}
\acro{bss}[BSS]{basic service set}
\acro{bssid}[BSSID]{\acl{bss} ID}
\acro{ccmp}[CCMP]{Counter Cipher Mode with Block Chaining Message Authentication Code Protocol}
\acro{cts}[CTS]{Clear to Send}
\acro{dhcp}[DHCP]{Dynamic Host Configuration Protocol}
\acro{dos}[DoS]{denial of service}
\acro{essid}[ESSID]{extended \acl{ssid}}
\acro{gid}[grID]{group ID}
\acro{ibss}[IBSS]{independent \acl{bss}}
\acro{icmp}[ICMP]{Internet Control Message Protocol}
\acro{iv}[IV]{initialization vector}
\acro{mimo}[MIMO]{multiple-input, multiple-output}
\acro{mac}[MAC]{Media Access Control}
\acro{mtu}[MTU]{Maximum Transferable Unit}
\acro{nat}[NAT]{Network Address Translation}
\acro{nd}[ND]{Neighbor Discovery}
\acro{nic}{network interface card}
\acro{ptw}[PTW]{Pychkine, Tews, Weinmann}
\acro{pxe}[PXE]{preboot execution environment}
\acro{ram}[RAM]{Random Access Memory}
\acro{rts}[RTS]{Request to Send}
\acro{rtt}[RTT]{round trip time}
\acro{ssh}[SSH]{secure shell}
\acro{ssid}[SSID]{Service Set ID}
\acro{sta}[STA]{station}
\acro{tcp}[TCP]{Transmission Control Protocol}
\acro{udp}[UDP]{User Datagram Protocol}
\acro{vap}[VAP]{virtual access point}
\acro{wan}[WAN]{wide area network}
\acro{wep}[WEP]{Wired Equivalent Privacy}
\acro{wmn}{wireless mesh node}
\acro{wnic}{wireless \acl{nic}}
\acro{wpa}[WPA]{Wi-Fi Protected Access}
\acro{wps}[WPS]{Wi-Fi Protected Setup}
\end{acronym}