### 2017-02-22 Lab_1/solutions/L1-12-2.tex
Added some more Lab1 solutions and all Lab1 traces

Minimum: 0.853ms
Maximum: 1.536ms
Average: 1.194

Faster: not via AP but direct