### 2017-02-23 Lab_1/solutions/L1-5-1.tex
Filling in parts of Lab_1, some questions remain about L1-5-1 and L1-6-2, L1-10-2 onward still to be done.
Former-commit-id: 338fad62136bf05ef12a7e88fdfaaee88eba11b4

\begin{itemize}

\item Neighbor Solicitation:  STA1 is trying to find STA2 by broadcasting these
\item Neighbor Advertisement: STA2 telling STA1 it's L2 address
\item Ping Request from STA1 to STA2
\item Ping Reply from STA2 to STA1
%\item More Neighbor Solicitations and Advertisements follow, refreshing the previous bindings. Interestingly, these have source and destination different.
\end{itemize}