### 2017-02-24 Lab_3/solutions/L3-3-1.tex

The resulting udp and tcp streams (with rts/cts) have a throughput that is consistently lower than their non-rts/cts counterparts. Adding the rts/cts mechanism does not really help because the theoretical datarate is hard to achieve under normal circumstances.

in this case it is no different and we see that the added overhead of rts/cts slows down the realised throughput of the stream even more.