### 2017-02-24 Lab_4/solutions/L4-1-2.tex

278: broadcasted probe request for the ssid wmn-1-A sent by STA1
279: the answer is from AP to STA1 and contains the ssid as well

in the association request <pkt id> the ssid is still listed. 
further communication such as association requests also contain the ssid

only broadcasts are now gone such as pkt 5 in the first L411

-- only authentication seen