### 2017-02-24 Lab_4/solutions/L4-1-1.tex

5 broadcast beacon frame

2528 probe 
2529 probe reply
2694 association request van sta1 naar ap waar ssid in staat