### 2017-02-24 Lab_4/solutions/L4-2-1.tex

in de eerdere kreeg authentication als reply succesful, nu is het unspecified failure
<pkt ids>