### 2017-02-24 Lab_4/solutions/L4-2-2.tex

Mac address spoof. either two mac addresses will be on the network causing problems and crashing the network or the spoofed mac address will not be present at the time granting the attacker access to the network.