### 2017-02-26 Lab_1/solutions/L1-10-2.tex
Finished Lab1 answers. Added .gitignore for latex files
Former-commit-id: db97bdc772bc07c8d7e33e5fe4f66942f69d6ec3

There are three main types of frames observed, but each can be divided into separate subtypes. 
\begin{itemize}
\item Management frames
\begin{itemize}

\item Acknowledgement (e.g. 9)

\item Probe Request (e.g. 15, for SSID UAntwerpen)
\item Beacon Frame (e.g. 29, for SSID UAntwerpen)

\item Probe response (e.g. 313, for SSID wmn-1-A)
\item Action (e.g. 1172)
\end{itemize}

\item Control frames
\begin{itemize}
\item Request-to-send (e.g. 1)
\item Clear-to-send (e.g. 7)
\item 802.11 Block Ack (e.g. 4)
\item Acknowledgement (e.g. 9)
\end{itemize}
\item Data frames
\begin{itemize}
\item QoS Null function (e.g. 47)
\item Null function (e.g. 1178)
\item Data (e.g. 1967, these are the Neighbor soliciations/advertisements and pings)
\end{itemize}
\end{itemize}
There were also some 802.11ac-specific frames such as the VHT NDP Announcement (e.g. 12)