### 2017-02-26 Lab_1/solutions/L1-10-3.tex

Now we do see 802.11 headers, and no more Ethernet headers. In monitor mode the driver or network stack no longer converts the headers to Ethernet. Each frame contains the radio information header with some network statistics. This seems to contain roughly the same info as the Radiotap header described below, and is also a pseudo-header. \\
What follows is the 802.11 header. The precise set of fields it contains depends on the (sub)type of the frame. The data frame headers seem to contain all possible fields: The frame control field containing a type + subtype byte, a duration field, source and destination MAC addresses of initial sender and final receiver, source and destination MAC address of current transmission, the AP's MAC, the STA's MAC, a fragment number, a sequence number and a frame check sequence. \\
Only the data frames also have a Logical Link Control header. It identifies used L2 protocols and the type of L3 data contained.