### 2017-02-26 Lab_1/solutions/L1-6-1.tex

STA1's trace file (traces/L1-5-1.STA1.pcap) shows duplicate ping requests (e.g. packets 3 and 4) and some duplicate neighbor advertisements (e.g. 10 and 11) and solicitations (e.g. 12 and 13).\\
According to the src MAC addresses, these are all packets sent by STA1.\\ \\
Similarly, for the STA2 dump (traces/L1-5-1.STA2.pcap), everything that is sent by STA2 is duplicated. \\
This includes the neighbor advertisement in packets 2 and 3; the ping replies in 5 and 6; the neighbor solicitation in 10 and 11.