### 2017-02-26 Lab_1/solutions/L1-10-4.tex
Finished Lab1 answers. Added .gitignore for latex files

This is the fully expanded radiotap header of packet 1977 (ping request).

\begin{lstlisting}
Radiotap Header v0, Length 26
    Header revision: 0
    Header pad: 0
    Header length: 26
    Present flags
        .... .... .... .... .... .... .... ...1 = TSFT: Present
        .... .... .... .... .... .... .... ..1. = Flags: Present
        .... .... .... .... .... .... .... .1.. = Rate: Present
        .... .... .... .... .... .... .... 1... = Channel: Present
        .... .... .... .... .... .... ...0 .... = FHSS: Absent
        .... .... .... .... .... .... ..1. .... = dBm Antenna Signal: Present
        .... .... .... .... .... .... .0.. .... = dBm Antenna Noise: Absent
        .... .... .... .... .... .... 0... .... = Lock Quality: Absent
        .... .... .... .... .... ...0 .... .... = TX Attenuation: Absent
        .... .... .... .... .... ..0. .... .... = dB TX Attenuation: Absent
        .... .... .... .... .... .0.. .... .... = dBm TX Power: Absent
        .... .... .... .... .... 1... .... .... = Antenna: Present
        .... .... .... .... ...0 .... .... .... = dB Antenna Signal: Absent
        .... .... .... .... ..0. .... .... .... = dB Antenna Noise: Absent
        .... .... .... .... .1.. .... .... .... = RX flags: Present
        .... .... .... .0.. .... .... .... .... = Channel+: Absent
        .... .... .... 0... .... .... .... .... = MCS information: Absent
        .... .... ...0 .... .... .... .... .... = A-MPDU Status: Absent
        .... .... ..0. .... .... .... .... .... = VHT information: Absent
        ...0 0000 00.. .... .... .... .... .... = Reserved: 0x00000000
        ..0. .... .... .... .... .... .... .... = Radiotap NS next: False
        .0.. .... .... .... .... .... .... .... = Vendor NS next: False
        0... .... .... .... .... .... .... .... = Ext: Absent
    MAC timestamp: 66921462
    Flags: 0x10
        .... ...0 = CFP: False
        .... ..0. = Preamble: Long
        .... .0.. = WEP: False
        .... 0... = Fragmentation: False
        ...1 .... = FCS at end: True
        ..0. .... = Data Pad: False
        .0.. .... = Bad FCS: False
        0... .... = Short GI: False
    Data Rate: 48,0 Mb/s
    Channel frequency: 5180 [A 36]
    Channel flags: 0x0140, Orthogonal Frequency-Division Multiplexing (OFDM), 5 GHz spectrum
        .... .... ...0 .... = Turbo: False
        .... .... ..0. .... = Complementary Code Keying (CCK): False
        .... .... .1.. .... = Orthogonal Frequency-Division Multiplexing (OFDM): True
        .... .... 0... .... = 2 GHz spectrum: False
        .... ...1 .... .... = 5 GHz spectrum: True
        .... ..0. .... .... = Passive: False
        .... .0.. .... .... = Dynamic CCK-OFDM: False
        .... 0... .... .... = Gaussian Frequency Shift Keying (GFSK): False
        ...0 .... .... .... = GSM (900MHz): False
        ..0. .... .... .... = Static Turbo: False
        .0.. .... .... .... = Half Rate Channel (10MHz Channel Width): False
        0... .... .... .... = Quarter Rate Channel (5MHz Channel Width): False
    SSI Signal: -50 dBm
    Antenna: 1
    RX flags: 0x0000
        .... .... .... .... .... ..0. = Bad PLCP: False
\end{lstlisting}
We'll only go over the actual statistics in the header and not every field. \\
There are some flags identifying the network's configuration in the Flags field (no WEP, fragmentation disabled, ...)\\
The data rate (48Mbps) and channel frequency (5180MHz, channel 36) are listed. \\
Then there are some channel flags identifying channel properties such as used multiplexing and modulating techniques. \\
Finally there is some info on signal strength and the used antenna.