### 2017-02-26 Lab_3/solutions/L3-2-1.tex
Added Lab3 answers and questions file

\includegraphics[width=0.5\textwidth]{traces/L3-2-1-tput.pdf}
\includegraphics[width=0.5\textwidth]{traces/L3-2-1-usage.pdf}
For low bitrates TCP performs better than UDP. Due to TCP's congestion control there are fewer collisions compared to UDP. For lower bitrates this makes TCP perform better than UDP. For higher bitrates the influence of TCP's additional overhead becomes larger and UDP is more performant than TCP again. \\ \\
For both TCP and UDP, throughput in this case is roughly half of the previous test. Each packet is transmitted twice (to the AP and from the AP), halving the total throughput between source and destination. \\ \\
 The effect of increasing bitrate on the usage is similar to the previous test.