### 2017-02-27 Lab_2/solutions/L2-6-2.tex
Working on Lab 2, filled in up to 4, half-assed 5 and 6: still has to be looked at for finishing. 802.11 (7-8) left as is for now because the results were inconclusive.
Former-commit-id: bd8cc1a2c8a31bdc0890d43ce4838e97a8e3ab36

Comparing the packets in 1611 (RTS) and 1612 (CTS), we immediately see that the RTS is targetted to the desired receiver while the CTS is broadcast. The RTS message is 6 bytes longer than the CTS message and the important differences are in the final field: for RTS it is 802.11 Request To Send and for CTS it is IEEE 802.11 Clear To Send.
The difference in size is because the L2 address of the transmitter of the CTS packet is not included in the CTS, while it is included for the RTS. Aside from that, not much differences can be observed: subtypes and such are different as expected.