### 2017-02-27 Lab_2/solutions/L2-3-1.tex
Working on Lab 2, filled in up to 4, half-assed 5 and 6: still has to be looked at for finishing. 802.11 (7-8) left as is for now because the results were inconclusive.
Former-commit-id: 37ac7684f000734feb181df1407cfca635225d9d [formerly 37ac7684f000734feb181df1407cfca635225d9d [formerly bd8cc1a2c8a31bdc0890d43ce4838e97a8e3ab36]]
Former-commit-id: 70f9d6ffe5623aee9e82549a39cc1f24d93efe24
Former-commit-id: 1f552c2cdc9567cf305dcbeca05b35bb858cfc6f

\begin{tabular}{l | l}
Selected channels & Observed channels in received beacons \\ \hline
1 & 1,4\\ \hline
2 & 1,4\\ \hline
3 & 1,4,6\\ \hline
4 & 4,6\\ \hline
5 & 4,6\\ \hline
6 & 4,6\\ \hline
7 & 6\\ \hline
8 & 4,6\\ \hline
9 & 11,4,6\\ \hline
10 & 11,4\\ \hline
11 & 11,4\\
\end{tabular}