### 2017-02-27 Lab_2/solutions/L2-5-1.tex

For this exercise, we will be comparing packet 2 from traces/L2-4-1.36.pcap and packet 2 from traces/L2-5-1.STA1.pcap.
For the purpose of comparing the WLAN management frames, I have included them below:

For the beacon frame captured in  exercise 4:
\begin{lstlisting}
IEEE 802.11 wireless LAN management frame
    Fixed parameters (12 bytes)
        Timestamp: 0x00000000082f503b
        Beacon Interval: 0,102400 [Seconds]
        Capabilities Information: 0x0001
            .... .... .... ...1 = ESS capabilities: Transmitter is an AP
            .... .... .... ..0. = IBSS status: Transmitter belongs to a BSS
            .... ..0. .... 00.. = CFP participation capabilities: No point coordinator at AP (0x0000)
            .... .... ...0 .... = Privacy: AP/STA cannot support WEP
            .... .... ..0. .... = Short Preamble: Not Allowed
            .... .... .0.. .... = PBCC: Not Allowed
            .... .... 0... .... = Channel Agility: Not in use
            .... ...0 .... .... = Spectrum Management: Not Implemented
            .... .0.. .... .... = Short Slot Time: Not in use
            .... 0... .... .... = Automatic Power Save Delivery: Not Implemented
            ...0 .... .... .... = Radio Measurement: Not Implemented
            ..0. .... .... .... = DSSS-OFDM: Not Allowed
            .0.. .... .... .... = Delayed Block Ack: Not Implemented
            0... .... .... .... = Immediate Block Ack: Not Implemented
    Tagged parameters (28 bytes)
        Tag: SSID parameter set: wmn-1-A
            Tag Number: SSID parameter set (0)
            Tag length: 7
            SSID: wmn-1-A
        Tag: Supported Rates 6(B), 9, 12(B), 18, 24(B), 36, 48, 54, [Mbit/sec]
            Tag Number: Supported Rates (1)
            Tag length: 8
            Supported Rates: 6(B) (0x8c)
            Supported Rates: 9 (0x12)
            Supported Rates: 12(B) (0x98)
            Supported Rates: 18 (0x24)
            Supported Rates: 24(B) (0xb0)
            Supported Rates: 36 (0x48)
            Supported Rates: 48 (0x60)
            Supported Rates: 54 (0x6c)
        Tag: DS Parameter set: Current Channel: 36
            Tag Number: DS Parameter set (3)
            Tag length: 1
            Current Channel: 36
        Tag: Traffic Indication Map (TIM): DTIM 1 of 0 bitmap
            Tag Number: Traffic Indication Map (TIM) (5)
            Tag length: 4
            DTIM count: 1
            DTIM period: 2
            Bitmap control: 0x00
                .... ...0 = Multicast: False
                0000 000. = Bitmap Offset: 0x00
            Partial Virtual Bitmap: 00
\end{lstlisting}
%
for the beacon frame captured in exercise 5:
\begin{lstlisting}
IEEE 802.11 wireless LAN management frame
    Fixed parameters (12 bytes)
        Timestamp: 0x000000000a53c0aa
        Beacon Interval: 0,102400 [Seconds]
        Capabilities Information: 0x0002
            .... .... .... ...0 = ESS capabilities: Transmitter is a STA
            .... .... .... ..1. = IBSS status: Transmitter belongs to an IBSS
            .... ..0. .... 00.. = CFP participation capabilities: Station is not CF-Pollable (0x0000)
            .... .... ...0 .... = Privacy: AP/STA cannot support WEP
            .... .... ..0. .... = Short Preamble: Not Allowed
            .... .... .0.. .... = PBCC: Not Allowed
            .... .... 0... .... = Channel Agility: Not in use
            .... ...0 .... .... = Spectrum Management: Not Implemented
            .... .0.. .... .... = Short Slot Time: Not in use
            .... 0... .... .... = Automatic Power Save Delivery: Not Implemented
            ...0 .... .... .... = Radio Measurement: Not Implemented
            ..0. .... .... .... = DSSS-OFDM: Not Allowed
            .0.. .... .... .... = Delayed Block Ack: Not Implemented
            0... .... .... .... = Immediate Block Ack: Not Implemented
    Tagged parameters (32 bytes)
        Tag: SSID parameter set: wmn-1-A
            Tag Number: SSID parameter set (0)
            Tag length: 7
            SSID: wmn-1-A
        Tag: Supported Rates 6(B), 9, 12(B), 18, 24(B), 36, 48, 54, [Mbit/sec]
            Tag Number: Supported Rates (1)
            Tag length: 8
            Supported Rates: 6(B) (0x8c)
            Supported Rates: 9 (0x12)
            Supported Rates: 12(B) (0x98)
            Supported Rates: 18 (0x24)
            Supported Rates: 24(B) (0xb0)
            Supported Rates: 36 (0x48)
            Supported Rates: 48 (0x60)
            Supported Rates: 54 (0x6c)
        Tag: IBSS Parameter set: ATIM window 0x0
            Tag Number: IBSS Parameter set (6)
            Tag length: 2
            Atim Windows: 0x0000
        Tag: Vendor Specific: Microsof: WMM/WME: Information Element
            Tag Number: Vendor Specific (221)
            Tag length: 7
            OUI: 00-50-f2
            Vendor Specific OUI Type: 2
            Type: WMM/WME (0x02)
            WME Subtype: Information Element (0)
            WME Version: 1
            WME QoS Info: 0x00
                0... .... = U-APSD: Disabled
                .... 0000 = Parameter Set Count: 0x00
                .000 .... = Reserved: 0x00
\end{lstlisting}
%
A first difference we observed is that the capabilities information beacon frame captured
in exercise 5 contains different capabilities information than the one captured in exercise 4:
the ex5 beacon frame shows that the sender of the beacon frame is a STA and that the network is
an ad-hoc network, while in exercise 4 the capabilities reflect that the transmitter is an AP
and the network is not ad-hoc.\\
For the tagged parameters section, the ex5 beacon is bigger. It contains some IBSS parameters (ATIM) and vendor specifici information about the ad-hoc network.
 In the ex4 beacon, the tagged parameters contain information about the channel and the Traffic Indication Map.