### 2017-02-27 Lab_2/solutions/L2-5-3.tex
Working on Lab 2, filled in up to 4, half-assed 5 and 6: still has to be looked at for finishing. 802.11 (7-8) left as is for now because the results were inconclusive.

The purpose of each of these frames:
\begin{itemize}
\item Request-To-Send packet
\item Clear-To-Send packet
\item 802.11 Block Ack
\item Acknowledgement, ie 2259
\item Probe Response ie 2601
\item Probe Request ie 2740
\end{itemize}