### 2017-03-01 Lab_4/solutions/L4-4-1.tex
Lab 4 exercises done, traces to come.

Difference : WEP parameters 4 bytes + 4 bytes that dont show up when you look at wep params in the pkt
IV en key index voor data, WEP ICV na data maar voor fcs