### 2017-03-01 Lab_2/solutions/L2-8-1.tex

No, the behaviour is largely the same: we again capture some icmp traffic but it is again not the complete conversation.
Because the stations are using the 36 channel (albeit with 40 MHz bands) and we are capturing on channel 36, we capture some icmp packets.
%
The 40 MHz wide bands indicate that the network has a "main" channel, in this case being 36, and extra space to be used if need be.
Because our icmp traffic is not enough to fill up the main channel, the extra space remains (largely) unused and we thus observe the same behaviour as in the previous exercise where we only capture part of the icmp exchange.