### 2017-03-02 Lab_3/solutions/L3-2-1.tex
Finished/improvied L3 answers, fixed L3-4 graph. Cleaned up L3 files
Former-commit-id: e2d9a5629b298fdd5f25bbca670651d596d6a449 [formerly a957ecc47cebd3f78a38d769264f36af02aa977c [formerly a957ecc47cebd3f78a38d769264f36af02aa977c [formerly fc51ba109bff8af870c143cfa139df67e8dd7573]]]
Former-commit-id: 9ad44cf59776ac9a2453877d9c225a3462773819
Former-commit-id: 98f291c9e89e1c8647b54ffef7ba3c97147b2df5
Former-commit-id: 00d79017118fbe7a6e9310e9051be078fe6d10fb

\includegraphics[width=0.5\textwidth]{traces/L3-2-1-tput.pdf}
\includegraphics[width=0.5\textwidth]{traces/L3-2-1-usage.pdf}
For low bitrates TCP performs better than UDP. Due to TCP's congestion control there are fewer collisions compared to UDP. With lower bitrates the attainable maximum throughput is lower. This means TCP's slow start will reach it faster. The shorter the slow start phase, the smaller its negative effect on throughput can be. For lower bitrates this makes TCP perform better than UDP. For higher bitrates the influence of TCP's additional overhead and negative effect of congestion control become larger and UDP is more performant than TCP again. \\ \\
For both TCP and UDP, throughput in this case is roughly half of the previous test. Each packet is transmitted twice (to the AP and from the AP), halving the total throughput between source and destination. \\ \\
 The effect of increasing bitrate on the usage is similar to the previous test.