### 2017-03-02 Lab_4/solutions/L4-1-1.tex

\begin{itemize}
\item PacketID example 5   , Beacon frame broadcast by the AP (or distributed amongst stations in an IBSS network) contains the SSID, in this case, wmn-1-A.
\item PacketID example 2528, Probe Request, in case of directed probe request contains the SSID of the network the station wants more information about. The other case is when no SSID is specified and "Broadcast" is used as SSID, in that case no SSID name is visible in this packet.
\item PacketID example 2529, Probe Response, response from AP containing capability information, supported data rates, etc., sent after receiving a probe request. The SSID is listed under tagged parameters and is equal to wmn-1-A.
\item PacketID example 2694, Association Request, part of the process where a station registers and connects to an AP. This first step contains the SSID with which the STA wants to connect and is sent to the corresponding AP.
\end{itemize}