### 2017-03-02 Lab_4/solutions/L4-1-2.tex

The SSID itself can still be seen in following packet / frame types, despite SSID hiding:
\begin{itemize}
\item PacketID 455  , Probe Request
\item PacketID 456  , Probe Response
\item PacketID 617  , Association Request
\end{itemize}
The beacon frames sent by our AP can no longer be used to determine the SSID but other traffic still works.
With the parameter being "ignore_broadcast_ssid", it was to be expected only the beacon frames would no longer contain the SSID.