### 2017-03-02 Lab_4/solutions/L4-4-1.tex
Writing Lab_4 report, finished up to and including WEP. Also limited the large tcpdump file as requested in the assignment.
Former-commit-id: 7fc75470399e852cd4643990a68d792c4b53d991

The difference in size between the decrypted encrypted packet with ID 434 in traces/L4-3-1.STA2.wep.pcap and the regular (non-WEP) packet with ID 353 in traces/L4-3-1.STA2.pcap is 8 bytes.
The extra 8 bytes that the decrypted encrypted packet has are listed in it's IEEE 802.11 Data header, as the WEP parameters. These WEP parameters include an initialisation vector
that is 3 bytes in length, a key index for 1 byte and also a WEP ICV that is 4 bytes long and is added after the encrypted data in the packet and before the frame check sequence.