### 2017-03-02 Lab_4/solutions/L4-2-2.tex

This security measure can be circumvented by spoofing our MAC-address to be among those listed in the accepted MAC addresses. \\ \\In the case that the original device
as well as the one we're currently spoofed as are present in the network, problems will arise in the network and things will break (the manner in which they break is
dependent on the type of APs and drivers used.). In the second case, when the original device is no longer present in the network
(this could be ensured by executing a deauthentication attack on the original device), the attacker will be able to access the network using the spoofed MAC address.