### 2017-03-02 Lab_2/solutions/L2-7-1.tex
Modified some L2 answers, cleaned up the L2 traces. Apart from 2-5-2 L2 is roughly finished
Former-commit-id: 147b063b2b4b0cc6af7827f13c3c03686ff0b711

We observed packets of both ping sessions, from the first session only the replies were caught on the monitor interface and for the second session, all ping packets were seen by the monitor.\\
%
Not seeing the ping requests of the first ping session could be due to failed demodulations at the monitor. Across multiple attemps almost all of STA2's ICMP packets were missed while the other STAs' were almost all captured. This could be related to STA1 and STA3 being close to each other and STA2 being farther away.
%
Failure to capture/demodulate STA2's packets may be related to the use of beamforming in 802.11n. This helps devices steer their RF energy right
to where the intended receiver is, and not to the monitor. This means the monitor will
get a worse signal to noise ratio, making is harder to demodulate packets sent at the highest data rates. We are however not sure whether the devices use/support beamforming.