### 2017-03-02 Lab_2/solutions/L2-5-1.tex
Modified some L2 answers, cleaned up the L2 traces. Apart from 2-5-2 L2 is roughly finished
Former-commit-id: e23d15b4d046db2d1a7efe0d385cd2063bea80e3 [formerly 667ac0bcf380efc65a8adcd7028db2ed790746ce [formerly 667ac0bcf380efc65a8adcd7028db2ed790746ce [formerly 147b063b2b4b0cc6af7827f13c3c03686ff0b711]]]
Former-commit-id: 9335664ea7e24b951bcf88594bcb3acc14c11ada
Former-commit-id: 0880c005726e10a7200c71b6dd62eef2a777fe49
Former-commit-id: 2e53566381fd089a1354c56a8389e71d68277fe5

For this exercise, we will be comparing packet 2 from traces/L2-4-1.36.pcap and packet 2 from traces/L2-5-1.STA1.pcap.
They are included below for convenience:

For the beacon frame captured in  exercise 4:
\begin{lstlisting}
IEEE 802.11 wireless LAN management frame
    Fixed parameters (12 bytes)
        Timestamp: 0x00000000082f503b
        Beacon Interval: 0,102400 [Seconds]
        Capabilities Information: 0x0001
            .... .... .... ...1 = ESS capabilities: Transmitter is an AP
            .... .... .... ..0. = IBSS status: Transmitter belongs to a BSS
            .... ..0. .... 00.. = CFP participation capabilities: No point coordinator at AP (0x0000)
            .... .... ...0 .... = Privacy: AP/STA cannot support WEP
            .... .... ..0. .... = Short Preamble: Not Allowed
            .... .... .0.. .... = PBCC: Not Allowed
            .... .... 0... .... = Channel Agility: Not in use
            .... ...0 .... .... = Spectrum Management: Not Implemented
            .... .0.. .... .... = Short Slot Time: Not in use
            .... 0... .... .... = Automatic Power Save Delivery: Not Implemented
            ...0 .... .... .... = Radio Measurement: Not Implemented
            ..0. .... .... .... = DSSS-OFDM: Not Allowed
            .0.. .... .... .... = Delayed Block Ack: Not Implemented
            0... .... .... .... = Immediate Block Ack: Not Implemented
    Tagged parameters (28 bytes)
        Tag: SSID parameter set: wmn-1-A
            Tag Number: SSID parameter set (0)
            Tag length: 7
            SSID: wmn-1-A
        Tag: Supported Rates 6(B), 9, 12(B), 18, 24(B), 36, 48, 54, [Mbit/sec]
            Tag Number: Supported Rates (1)
            Tag length: 8
            Supported Rates: 6(B) (0x8c)
            Supported Rates: 9 (0x12)
            Supported Rates: 12(B) (0x98)
            Supported Rates: 18 (0x24)
            Supported Rates: 24(B) (0xb0)
            Supported Rates: 36 (0x48)
            Supported Rates: 48 (0x60)
            Supported Rates: 54 (0x6c)
        Tag: DS Parameter set: Current Channel: 36
            Tag Number: DS Parameter set (3)
            Tag length: 1
            Current Channel: 36
        Tag: Traffic Indication Map (TIM): DTIM 1 of 0 bitmap
            Tag Number: Traffic Indication Map (TIM) (5)
            Tag length: 4
            DTIM count: 1
            DTIM period: 2
            Bitmap control: 0x00
                .... ...0 = Multicast: False
                0000 000. = Bitmap Offset: 0x00
            Partial Virtual Bitmap: 00
\end{lstlisting}
%
For the beacon frame captured in exercise 5:
\begin{lstlisting}
IEEE 802.11 wireless LAN management frame
    Fixed parameters (12 bytes)
        Timestamp: 0x000000000a53c0aa
        Beacon Interval: 0,102400 [Seconds]
        Capabilities Information: 0x0002
            .... .... .... ...0 = ESS capabilities: Transmitter is a STA
            .... .... .... ..1. = IBSS status: Transmitter belongs to an IBSS
            .... ..0. .... 00.. = CFP participation capabilities: Station is not CF-Pollable (0x0000)
            .... .... ...0 .... = Privacy: AP/STA cannot support WEP
            .... .... ..0. .... = Short Preamble: Not Allowed
            .... .... .0.. .... = PBCC: Not Allowed
            .... .... 0... .... = Channel Agility: Not in use
            .... ...0 .... .... = Spectrum Management: Not Implemented
            .... .0.. .... .... = Short Slot Time: Not in use
            .... 0... .... .... = Automatic Power Save Delivery: Not Implemented
            ...0 .... .... .... = Radio Measurement: Not Implemented
            ..0. .... .... .... = DSSS-OFDM: Not Allowed
            .0.. .... .... .... = Delayed Block Ack: Not Implemented
            0... .... .... .... = Immediate Block Ack: Not Implemented
    Tagged parameters (32 bytes)
        Tag: SSID parameter set: wmn-1-A
            Tag Number: SSID parameter set (0)
            Tag length: 7
            SSID: wmn-1-A
        Tag: Supported Rates 6(B), 9, 12(B), 18, 24(B), 36, 48, 54, [Mbit/sec]
            Tag Number: Supported Rates (1)
            Tag length: 8
            Supported Rates: 6(B) (0x8c)
            Supported Rates: 9 (0x12)
            Supported Rates: 12(B) (0x98)
            Supported Rates: 18 (0x24)
            Supported Rates: 24(B) (0xb0)
            Supported Rates: 36 (0x48)
            Supported Rates: 48 (0x60)
            Supported Rates: 54 (0x6c)
        Tag: IBSS Parameter set: ATIM window 0x0
            Tag Number: IBSS Parameter set (6)
            Tag length: 2
            Atim Windows: 0x0000
        Tag: Vendor Specific: Microsof: WMM/WME: Information Element
            Tag Number: Vendor Specific (221)
            Tag length: 7
            OUI: 00-50-f2
            Vendor Specific OUI Type: 2
            Type: WMM/WME (0x02)
            WME Subtype: Information Element (0)
            WME Version: 1
            WME QoS Info: 0x00
                0... .... = U-APSD: Disabled
                .... 0000 = Parameter Set Count: 0x00
                .000 .... = Reserved: 0x00
\end{lstlisting}
%
A first difference we observed is in the capabilities information parameters:
the ad-hoc beacon frame shows that the sender of the beacon frame is a STA and that the network is
an ad-hoc network, while the AP's capabilities reflect that the transmitter is an AP
and the network is not in an IBSS (i.e. not ad-hoc).\\
For the tagged parameters section, the two contain different parameter sets, making the ad-hoc one 4 bytes longer. The ad-hoc frame contains an IBSS parameter set with the Announcement Traffic Indication Message (ATIM) window and vendor specificic information parameters concerning Quality of Service.
 In the AP beacon, the tagged parameter sets contain information about the channel and the Traffic Indication Map.