### 2017-03-03 Lab_4/solutions/L4-2-2.tex
Changed some previous L4 answers, added L4-5 anwers with new trace file, cleaned up L4 traces
Former-commit-id: 5511c468bc01f5a2f070aae601f4fa113eea62a9 [formerly 44bdf5977a2e43760e4e50d86b41a27e37a221e3 [formerly cd8d4c681d2b595cbdffce27a6fdc369899f8747] [formerly cd8d4c681d2b595cbdffce27a6fdc369899f8747 [formerly d095ed18bcc71f5876daaae0cd962c7a5891cf27]]]
Former-commit-id: 0377acbb4cc17645053cd1131ef969e69a56490a
Former-commit-id: 6eda004ed558a2a57893992f193415d08eff9107
Former-commit-id: fcb4274de68a0f9471ff8ca397eafd6473101ed9

An attacker can easily circumvent this security measure by spoofing their MAC-address to be among those listed in the accepted MAC addresses. The easiest way to find such an address is to monitor the network and obtain MAC addresses of currently connected devices.  \\ \\In the case that the attacker uses a MAC address for which the genuine device is also connected, problems may arise in the network. MAC addresses are supposed to be unique within a LAN. The precise issues to arise depend on network configuration and the specific devices used. In a wired LAN one of the two devices would probably become inaccissble due to their entry being overwritten in the MAC table of the switch. This obviously cannot occur in a wireless LAN.