### 2017-03-03 Lab_4/solutions/L4-1-1.tex
Changed some previous L4 answers, added L4-5 anwers with new trace file, cleaned up L4 traces
Former-commit-id: d095ed18bcc71f5876daaae0cd962c7a5891cf27

\begin{itemize}
\item Beacon frame (packet 5): broadcast by the AP (or distributed amongst stations in an IBSS network) contains the SSID, in this case, wmn-1-A.
\item Probe Request (packet 2528): the directed probe request contains the SSID of the network the station wants more information about. There are also broadcast probe requests which do not contain an SSID.
\item Probe Response (packet 2529): response from AP containing capability information, supported data rates, etc., sent after receiving a probe request. The SSID is listed under tagged parameters.
\item Association Request (packet 2694): part of the process where a station registers and connects to an AP. This first step contains the SSID with which the STA wants to connect and is sent to the corresponding AP.
\end{itemize}