### 2017-03-03 Lab_4/solutions/L4-4-1.tex

The difference in size between the encrypted packet with ID 1415 in traces/L4-3-1.STA2.wep.pcap and the regular (non-WEP) packet with ID 1077 in traces/L4-3-1.STA2.pcap is 8 bytes (154 vs 146).
The extra 8 bytes are the WEP parameters in the IEEE 802.11 Data header. These WEP parameters are an initialization vector (3 bytes), a key index (1 byte) and a WEP Integrity Check Value (4 bytes). This ICV is inserted between the encrypted data and the frame check sequence.