### 2017-03-03 Lab_4/solutions/L4-6-1.tex
report on DoS (after finally understanding the whole spoof-everything concept) completed. Wrote report using current L4-6-1.pcap, the result of the second experiment and stored old result in L4-6-1.old_poging1.pcap.

The deassociation attack caused the ping session to be interrupted. After 5 seconds, the station was able to reassociate with the AP and resume the ping traffic.
Repeating the attack some time later again caused a 5 second interruption in ping traffic. \\ \\A second attempt at the exercise showed a 22 second delay and not a 5 second delay.
%
We found that reestablishment of the connection is a function of how aggressively the client attempts to reauthenticate and any higher-level timeouts or backoffs and varying network quality
that may suppress the demand for communication and conclude that that is what leads to the difference in the time needed to be able to resume the ping session.\\ \\
% %source: http://users.csc.calpoly.edu/~bellardo/pubs/usenix-sec03-80211dos-html/aio.html -- deauthentication header
The main thing to note for the ping session is that it was interrupted for a short time but when the attack ceased, the ping session resumed after the STA reassociated with the AP.