### 2017-03-05 Lab_4/solutions/L4-6-1.tex
Modified L4-6 solutions

The disassociation/deauthentication attack caused the ping session to be interrupted. After 5 seconds, the station was able to reestablish connection with the AP and resume the ping traffic.
Repeating the attack some time later again caused a  5 second interruption in ping traffic.  \\ \\
The main thing to note for the ping session is that it was interrupted for a short time but when the attack ceased, the ping session resumed quickly after the STA reassociated with the AP.