### 2017-03-05 Lab_4/solutions/L4-2-2.tex
Review of Lab_4, added refs to tracefiles where I saw they were missing. Fixed some typo's.

An attacker can easily circumvent this security measure by spoofing their MAC-address to be among those listed in the accepted MAC addresses. The easiest way to find such an address is to monitor the network and obtain MAC addresses of currently connected devices.  \\ \\In the case that the attacker uses a MAC address for which the genuine device is also connected, problems may arise in the network. MAC addresses are supposed to be unique within a LAN. The precise issues to arise depend on network configuration and the specific devices used. In a wired LAN one of the two devices would probably become inaccessible due to their entry being overwritten in the MAC table of the switch. This obviously cannot occur in a wireless LAN.