### 2017-03-05 Lab_4/solutions/L4-1-2.tex
Review of Lab_4, added refs to tracefiles where I saw they were missing. Fixed some typo's.
Former-commit-id: 180a98254dcf255543ff72f101deac34a2312d85

The SSID itself can still be seen in following packet / frame types in our dump file traces/L4-1-2.STA2.pcap, despite SSID hiding:
\begin{itemize}
\item Probe Request (packet 455)
\item Probe Response (packet 456)
\item Association Request (packet 617)
\end{itemize}
The SSID was only removed from the broadcast beacon frames, the others are unchanged. Note that the beacon frames are still sent. In Wireshark's packet info field the SSID is shown as "Broadcast" but further inspection of the packet shows the SSID field is simply empty.
With the parameter being "ignore\_broadcast\_ssid", it was to be expected only the beacon frames would no longer contain the SSID.