### 2017-03-05 Lab_4/solutions/L4-2-1.tex
Small changes to L4
Former-commit-id: 51d0f2b3db5e9158fedd338386d5e2e5972869e0 [formerly 2b5f1f1c0fb6825ae69dc645d22b140faa770073 [formerly 2b5f1f1c0fb6825ae69dc645d22b140faa770073 [formerly 85572e302b2054e9df8addb61e1b05a242c19184]]]
Former-commit-id: ecd77de1f9afec5cae939969c855182e2b9d2389
Former-commit-id: 52891089232b5d215ced4a8b3ee8043519173568
Former-commit-id: 21c20f630ade97911bbb9fd85a85221c0011d317

In the first dump (traces/L4-1-1.STA2.pcap), when the station sent an authentication frame (packet 2666) the AP replied with an authentication frame (packet 2668) indicating the station was succesfully authenticated (status code 0: successful).\\
In the second case (found in traces/L4-2-1.STA2.pcap), with MAC filtering enabled and configured to not allow STA1's wlan0 to connect,
the authentication frame (packet 464) sent by STA1 is met with an authentication frame (packet 466)
with status code 1. This indicates an unspecified failure. STA1 does not attempt an association as authentication failed. Note that STA1 is not informed that authentication failed because it was blacklisted. There is no reason to let a potential attacker know why exactly they could not connect.