### 2017-03-15 Lab_1/solutions/L1-5-1.tex
Final changes to lab1 report, updated zip file to reflect changes.

Looking at trace file traces/L1-5-1.AP.pcap, we observe the following message types:
\begin{itemize}
\item Neighbor Solicitation (ie packet nr 1):  STA1 is trying to find STA2 by broadcasting these
\item Neighbor Advertisement (ie packet nr 3): STA2 telling STA1 its L2 address
\item Ping Request (ie packet nr 4) from STA1 to STA2
\item Ping Reply (ie packet nr 5) from STA2 to STA1
%\item More Neighbor Solicitations and Advertisements follow, refreshing the previous bindings. Interestingly, these have different source and destination addresses.
\end{itemize}