### 2017-03-15 Lab_2/solutions/L2-3-1.tex
Adding some trace file names, edited L2-3-1 to reflect channel 4 being detected everywhere.

\begin{tabular}{l | l}
Selected channels & Observed channels in received beacons \\ \hline
1 & 1,4\\ \hline
2 & 1,4\\ \hline
3 & 1,4,6\\ \hline
4 & 4,6\\ \hline
5 & 4,6\\ \hline
6 & 4,6\\ \hline
7 & 6\\ \hline
8 & 4,6\\ \hline
9 & 11,4,6\\ \hline
10 & 11,4\\ \hline
11 & 11,4\\
\end{tabular}
Interestingly, channel 4 beacons are observed in every channel from 1 to 11, though only from wmn-1-A.\\
Other SSIDs on channel 4 are only observed when we monitor channel 4 and neighbouring channels. We believe this is due to the wmn-1-A AP being fairly close to our monitor.