### 2017-03-19 Lab_4/solutions/L4-4-1.tex
Minor fixes in L2-3-4 and added archives for those
Former-commit-id: 8e593e34ab95cfd21c878b5a66e993400c632762 [formerly 2b0ec8ef5fab7d8e2c62aaa9cd61650e387c6a71 [formerly dfd7b92f95c615fda60ea23507f0bc2517c4f04a] [formerly dfd7b92f95c615fda60ea23507f0bc2517c4f04a [formerly bb1a3edab162e967665c3611fdac1833ab1311f2]]]
Former-commit-id: 63eadee2fee1c67a9cfeb5f7adb8c8283e6c639e
Former-commit-id: 56b7edbcf2d86fa0d25247e2ac3cb001141d7488
Former-commit-id: 66c32bcb37571a831240aa9bf463085e19dc8c2d

The difference in size between the encrypted packet with ID 1415 in traces/L4-3-1.STA2.wep.pcap and the regular (non-WEP) packet with ID 1077 in traces/L4-3-1.STA2.pcap is 8 bytes (154 vs 146).
The extra 8 bytes are the WEP parameters in the IEEE 802.11 Data header. These WEP parameters are an initialization vector (3 bytes), a key index (1 byte) and a WEP Integrity Check Value (4 bytes). The IV and key index are right before the encrypted data, while the ICV is inserted between the encrypted data and the frame check sequence.